Using traffic analysis to identify The Second Generation Onion Router

### 2011-08-11 ieee-conference-paper/paper.tex
Some tweaking with figure positions and tables, 6 pages, slightly larger histograms

@@

To reduce the affect of confounding variables, all phases conducted the same
simulations, on an isolated test network and were conducted within a virtual
machine snapshot which was periodically rolled back to a clean, known state.
The experiment yielded three sets of sample data:

@@

The data captured was in the form of 1MB capture files which were recombined
with mergecap and processed by NetAI \parencite{swinbut:2006fk} to produce ARFF
format files for use by Weka \parencite{Hall:2009p7662,Bouckaert:2010we}. NetAI
identifies flows inside the capture files and produces a number of statistics
to be used as attributes for classification.

All of the algorithms chosen were able to successfully classify HTTPS and HTTP
over Tor traffic with accuracy in excess of 90\%, with the Adaboost algorithm
unable to classify HTTPS over Tor. The best performing algorithm random forest
was able to classify HTTP over Tor with 93.7\% accuracy and a false positive
rate of 3.7\%. HTTPS over Tor was more easily identified with a 97.7\% accuracy
and low false positive rate of 0.3\%.

@@

With some relative success classifying Tor traffic using unsupervised machine
learning techniques, the captures were examined for a heuristics based
classification approach. When a histogram of packet sizes for each of the
packet traces is produced, a clear distinction can between seen between the
sample sets:

\begin{figure}[h]
  \centering\includegraphics[width=\linewidth]{https_hist}
  \caption{Histogram of packet size for HTTPS traffic}
  \label{https_hist}
\end{figure}

\begin{figure}[h]
  \centering\includegraphics[width=\linewidth]{http_tor_hist}
  \caption{Histogram of packet size for HTTP traffic over Tor}
  \label{http_tor_hist}
\end{figure}

\begin{figure}[h]
  \centering\includegraphics[width=\linewidth]{https_tor_hist}
  \caption{Histogram of packet size for HTTPS traffic over Tor}
  \label{https_tor_hist}
\end{figure}

Investigating more closely, by examining individual packet traces some
significant patterns begin to appear. Table \ref{table:packet_sizes} shows the
packet sizes seen for the first 10 packets of a 1\% sample of sessions in each
phase of data captured. The first few packets contain 0 bytes of data, these
are the typical SYN, SYN/ACK and ACK flagged packets that are used to initiate
a TCP/IP connection. Following that only two packet sizes are seen for HTTPS
connections. For Tor sessions a range of packet sizes from 131 to 152. This
knowledge can be used to build a very rudimentary classification algorithm to
identify Tor traffic, as seen in the pseudo-code below:

\begin{lstlisting}[language=ruby]
if packet[3] > 130 && packet[3] <  153 && packet[5] > 913 && packet[5] < 937
  is_tor = true
else
  is_tor = false
end
\end{lstlisting}

Applying this algorithm to a 10\% sample of each of the packet traces yields
the results as seen in table \ref{table:heuristic-results}.

\begin{table}[h]
  \begin{tabular*}{\linewidth}{lr}
    \toprule
    Protocol & Identified as Tor\\
    \midrule
    HTTPS & 0.0000\% \\
    HTTP over Tor & 0.9903\% \\
    HTTPS over Tor & 0.9826\% \\
    \bottomrule
  \end{tabular*}
  \caption{Results}
  \label{table:heuristic-results}
\end{table}

@@

>>>>

\begin{figure}[H]
  \centering\includegraphics[width=\linewidth]{tor-network-diagram}
  \caption{The Tor Network}
  \label{tor-network}
\end{figure}

@@

To reduce the affect of confounding variables, all phases conducted the same
simulations, on an isolated test network and were conducted within a virtual
machine snapshot which was periodically rolled back to a clean, known state.
The experiment yielded three sets of, a summary is included in table
\ref{table:datasets}.

@@

The data captured was in the form of 1MB capture files which were recombined
with mergecap \parencite{Renfro:fk} and processed by NetAI
\parencite{swinbut:2006fk} to produce ARFF format files for use by Weka
\parencite{Hall:2009p7662,Bouckaert:2010we}. NetAI identifies flows inside the
capture files and produces a number of statistics to be used as attributes for
classification.

All of the algorithms chosen were able to successfully classify HTTPS and HTTP
over Tor traffic with accuracy in excess of 90\%, with the Adaboost algorithm
unable to classify HTTPS over Tor. The best performing algorithm random forest
was able to classify HTTP over Tor with 93.7\% accuracy and a false positive
rate of 3.7\%. HTTPS over Tor was more easily identified with a 97.7\% accuracy
and low false positive rate of 0.3\%. A summary of the results obtained by
these machine learning classifiers is available in Table \ref{table:results}.

@@

With some relative success classifying Tor traffic using unsupervised machine
learning techniques, the captures were examined for a heuristics based
classification approach. When a histogram of packet sizes for each of the
packet traces is produced, a clear distinction can between seen between the
sample sets as seen in Figures \ref{https_hist}, \ref{http_tor_hist} and
\ref{https_tor_hist}.

\begin{figure*}
  \centering\includegraphics[width=0.6\linewidth]{https_hist}
  \caption{Histogram of packet size for HTTPS traffic}
  \label{https_hist}
\end{figure*}

\begin{figure*}
  \centering\includegraphics[width=0.6\linewidth]{http_tor_hist}
  \caption{Histogram of packet size for HTTP traffic over Tor}
  \label{http_tor_hist}
\end{figure*}

\begin{figure*}
  \centering\includegraphics[width=0.6\linewidth]{https_tor_hist}
  \caption{Histogram of packet size for HTTPS traffic over Tor}
  \label{https_tor_hist}
\end{figure*}

Investigating more closely, by examining individual packet traces some
significant patterns begin to appear. Table \ref{table:packet_sizes} shows the
packet sizes seen for the first 6 packets of a 1\% sample of sessions in each
phase of data captured. The first few packets contain 0 bytes of data, these
are the typical SYN, SYN/ACK and ACK flagged packets that are used to initiate
a TCP/IP connection. Following that only two packet sizes are seen for HTTPS
connections. For Tor sessions a range of packet sizes from 131 to 152.

This knowledge can be used to build a rudimentary classification algorithm to
identify Tor traffic, the pseudo code for an example algorithm is included in
Figure \ref{psuedo_matcher}. Applying this algorithm to the complete traces
yields the results as seen in Table \ref{table:heuristic-results}.

\begin{table}[H]
  \begin{tabular*}{\linewidth}{lr}
    \toprule
    Protocol & Identified as Tor\\
    \midrule
    HTTPS & 1.06\%\\
    HTTP over Tor & 98.13\%\\
    HTTPS over Tor & 97.54\%\\
    \bottomrule
  \end{tabular*}
  \caption{Results from Heuristic Classifier}
  \label{table:heuristic-results}
\end{table}

@@

\begin{figure}
  \begin{lstlisting}[language=ruby]
    if  packet[3] > 130 and
        packet[3] <  153 and
        packet[5] > 913 and
        packet[5] < 937 then
      is_tor = true
    else
      is_tor = false
    end
  \end{lstlisting}
  \caption{Pseudo Code for Matching a Tor Session}
  \label{psuedo_matcher}
\end{figure}
